Locally Conservative Discontinuous Petrov-Galerkin Finite Elements for Fluid Problems

### 2013-12-04 docs/Truman/LocalConservation/paper/paper.tex
Fixed some equation parenthesis in LC paper

Classical $h$-interpolation interpolation error estimate for the lowest error
Raviart-Thomas elements and continuity of operator $R$ imply the stability estimate:
\begin{equation}
\begin{array}{lll}
|| \bfpsi_h || & \leq || \bfpsi_h - \bfpsi || + || \bfpsi ||\\[8pt]
&\leq C h || \bfpsi ||_{H^1} +  || \bfpsi || \\[8pt]
& \leq C || \text{div} \bfpsi || = C (\sum_K \mu(K) \lambda_K^2)^{1/2}\,.
\end{array}
\label{eq:stab}
\end{equation}
Above, $C$ is a generic, mesh independent constant incorporating constant from
the interpolation error estimate and continuity constant of $R$.
Let $\hat{t}$ be the trace of $\bfpsi_h$. We have then,
\begin{equation}
\sup_{\hat{t} \in H^{-1/2}(\Gamma_h)} \frac{|  \sum_K \lambda_K \langle
\hat{t},1_K \rangle_{\ptl K} |}{|| \hat{t} ||_{H^{-1/2}(\Gamma_h)}}
\geq \frac{| \sum_K \lambda_K \int_K \text{div} \bfpsi_h \: 1_K  |}
{|| \bfpsi_h ||_{H(\text{div},\Omega)}}
\geq \frac{1}{C} (\sum_K \mu(K) \lambda_K^2)^{1/2}\,,
\end{equation}
where $C$ is the constant from stability estimate~(\ref{eq:stab}).

@@

Recall the line of analysis leading to the construction of robust
test norms allowing us to bound the $L^2$ error of the field variables by the
energy error, Equation~(\ref{eq:robustBound}). With robust test norms, we have
\begin{equation}
\begin{array}{ll}
   \LRp{|| u - u_h ||^2+ || \bfsig - \bfsig_h ||^2}^{\frac{1}{2}}
& \lesssim || (u - u_h, \bfsig - \bfsig_h, \hat{u} - \hat{u}_h, \hat{t} - \hat{t}_h ||_E \\[8pt]
&= \inf_{(w_h,\bfvarsigma_h,\hat{w},\hat{r}_h)}
|| (u - w_h, \bfsig - \bfvarsigma_h, \hat{u} - \hat{w}_h, \hat{t} - \hat{r}_h ||_E\,.
\end{array}
\end{equation}
The last equality follows from the fact that DPG method delivers the best approximation
error in the energy norm (minimizes the residual). This is no longer true for the
restricted version. So, can we claim robustness in the sense of the inequality above
for the restricted version as well?

@@

Let us start with the inf-sup condition,
\begin{equation}
\sup_{\hat{t}} \frac{| \sum_K \lambda_K \langle \hat{t},1_K \rangle |}{|| \hat{t} ||_E}
\geq  \beta \LRp{\sum_K \mu(K) \lambda_K^2}^{1/2}\,.
\end{equation}
As in the previous analysis, we select for $\hat{t}$ the trace of Raviart-Thomas
interpolant $\bfpsi_h$ of $\bfpsi = R (\sum_K \lambda_K 1_K)$ where $R$ is the right-inverse
of the divergence operator constructed by Costabel and McIntosh. The only change compared
with the previous analysis, is the evaluation of norm of $\hat{t}_h$. For this, we need
to solve the local problems:
\begin{align}
((v,\bftau), (\delta v, \delta \bftau)_V &= \langle \hat{t}, \delta v \rangle_{\ptl K}
= \int_K \text{div} \bfpsi_h \, \delta v = \int_K \text{div} \bfpsi \, \delta v
\nonumber\\
&= \int_K \lambda_K \delta v
= \lambda_K (1_K,\delta v)_K  \quad \forall \delta v \in
H^1(K)\,\forall\delta\bftau\in\bfH(\text{div},K)\,.
\label{eq:local}
\end{align}
We need then an upper bound of the energy norm of $(v_h,\bftau_h)$:
$$
\LRp{\sum_K || (v,\bftau) ||_V^2}^{1/2}\,.
$$
Substituting $(v,\bftau)$ for $(\delta v,\delta\bftau)$ in~(\ref{eq:local}), we get,
\begin{equation}
|| (v,\bftau) ||_V^2 = \lambda_K (1_K, v_K)\,.
\end{equation}
If we have a robust stability estimate:
\begin{equation}
(1_K, v_K) \leq C \mu(K)^{1/2} || (v,\bftau) ||_K
\label{eq:robustEst1}
\end{equation}
(i.e. constant $C$ is independent of $\epsilon$), then
\begin{equation}
||  (v,\bftau) ||_V \leq C \mu(K)^{1/2} | \lambda_K |
\end{equation}
and, eventually as needed,
\begin{equation}
\sum_K || (v,\bftau) ||_V^2  \leq C^2 \sum_K \mu(K) \lambda_K^2\,,
\end{equation}
which leads to the robust estimate of inf-sup constant $\beta$. For example, it is sufficient if
\begin{equation}
|| v ||_{L^2(K)} \leq || (v,\bftau) ||_V \, .
\label{eq:robust_est1}
\end{equation}
Notice that the stability analysis with the energy norm was, in a sense, easier than
with the quotient norm. Only the divergence of the interpolant $\bfpsi_h$ enters~(\ref{eq:local})
and it coincides with the divergence of $\bfpsi$.

We arrive at a similar situation in the continuity estimate of
$$
\sum_K \lambda_K \langle \hat{t}, 1_K \rangle\,.
$$
Testing with $(1_K,{\bf 0})$ in the local problem~(\ref{eq:local_problem}), we obtain,
\begin{equation}
((v,\bftau),(1_K,{\bf 0}))_V = \langle \hat{t}, 1_K \rangle_{\ptl K}\,.
\label{eq:local2}
\end{equation}
If we have a robust estimate,
\begin{equation}
| ((v,\bftau),(1_K,{\bf 0}))_V | \leq C \mu(K)^{1/2} \, || (v,\bftau) ||_V\,,
\label{eq:robust_est2}
\end{equation}
then
\begin{equation}
| \sum_K \lambda_K \langle \hat{t}, 1_K \rangle | \leq C (\sum_K \mu(K) \lambda_K^2)^{1/2}
\, (\sum_K || (v,\bftau) ||_V^2)^{1/2}
= C (\sum_K \mu(K) \lambda_K^2)^{1/2}  || \hat{t} ||_E\,,
\end{equation}
as needed.

For instance, condition~(\ref{eq:robust_est2}) will be satisfied if the test inner product
in~(\ref{eq:local2}) reduces to the $L^2$ term only,
\begin{equation}
((v,\bftau),(1_K,{\bf 0}))_V = (v,1_K)_{L^2(K)} \: .
\end{equation}
With the robust stability and continuity constants for the mixed problem, the energy error
of solution $(u,\bfsig,\hat{u},\hat{t})$ (and Lagrange multipliers $\lambda_K$ as well)
is bounded robustly by the {\em best approximation error} of  $(u,\bfsig,\hat{u},\hat{t})$
measured in the energy norm. We arrive thus at the same situation as in the standard
DPG method.

@@

We now proceed to show that the robust norms we are using satisfy this
requirement.
Consider the inner product from Equation~(\ref{eq:localSolve}), with
$\alpha=1$. We wish to
verify condition (\ref{eq:robustEst1}) with the norm derived from this inner
product on the right hand side. By Cauchy-Schwarz
\begin{equation}
   \int_K v\cdot
   1\leq\mu(K)^{1/2}\norm{v}_{L^2(K)}\leq\mu(K)^{1/2}\norm{(v,\btau)}_K\,,
   \label{eq:robustProof}
\end{equation}
where $\norm{(v,\btau)}_K$ is the norm derived from the inner product.
Condition~(\ref{eq:robust_est2}) comes out the same since
$$|(
(v,\btau),(1_K,{\bf0}))| =
\sum_K|(1_K,v_K)|\leq\sum_K\mu(K)^{1/2}\norm{v,\btau}_K
$$
element-wise.

Now we wish the perform the same analysis for the modified inner product in
Equation~(\ref{eq:localSolveMod}). In this case,
condition~(\ref{eq:robustEst1}) follows even more naturally as
\begin{equation}
   \int_K v\cdot1\leq\mu(K)^{1/2}\frac{1}{\mu(K)^{1/2}}\LRl{\int_Kv}
   \leq\norm{(v,\btau)}_K\,,
   \label{eq:robustProof2}
\end{equation}
where $\norm{(v,\btau)}$ now refers to the norm generated by inner product
(\ref{eq:localSolveMod}). Condition~(\ref{eq:robust_est2}) follows by the same
reasoning.

>>>>

Classical $h$-interpolation interpolation error estimate for the lowest error
Raviart-Thomas elements and continuity of operator $R$ imply the stability estimate:
\begin{equation}
\begin{array}{lll}
|| \bfpsi_h || & \leq || \bfpsi_h - \bfpsi || + || \bfpsi ||\\[8pt]
&\leq C h || \bfpsi ||_{H^1} +  || \bfpsi || \\[8pt]
& \leq C || \text{div} \bfpsi || = C (\sum_K \mu(K) \lambda_K^2)^{1/2}\,.
\end{array}
\label{eq:stab}
\end{equation}
Above, $C$ is a generic, mesh independent constant incorporating constant from
the interpolation error estimate and continuity constant of $R$.
Let $\hat{t}$ be the trace of $\bfpsi_h$. We have then,
\begin{equation}
\sup_{\hat{t} \in H^{-1/2}(\Gamma_h)} \frac{|  \sum_K \lambda_K \langle
\hat{t},1_K \rangle_{\ptl K} |}{|| \hat{t} ||_{H^{-1/2}(\Gamma_h)}}
\geq \frac{| \sum_K \lambda_K \int_K \text{div} \bfpsi_h \: 1_K  |}
{|| \bfpsi_h ||_{H(\text{div},\Omega)}}
\geq \frac{1}{C} (\sum_K \mu(K) \lambda_K^2)^{1/2}\,,
\end{equation}
where $C$ is the constant from stability estimate~\eqref{eq:stab}.

@@

Recall the line of analysis leading to the construction of robust
test norms allowing us to bound the $L^2$ error of the field variables by the
energy error, \eqref{eq:robustBound}. With robust test norms, we have
\begin{equation}
\begin{array}{ll}
   \LRp{|| u - u_h ||^2+ || \bfsig - \bfsig_h ||^2}^{\frac{1}{2}}
& \lesssim || (u - u_h, \bfsig - \bfsig_h, \hat{u} - \hat{u}_h, \hat{t} - \hat{t}_h ||_E \\[8pt]
&= \inf_{(w_h,\bfvarsigma_h,\hat{w},\hat{r}_h)}
|| (u - w_h, \bfsig - \bfvarsigma_h, \hat{u} - \hat{w}_h, \hat{t} - \hat{r}_h ||_E\,.
\end{array}
\end{equation}
The last equality follows from the fact that DPG method delivers the best approximation
error in the energy norm (minimizes the residual). This is no longer true for the
restricted version. So, can we claim robustness in the sense of the inequality above
for the restricted version as well?

@@

Let us start with the inf-sup condition,
\begin{equation}
\sup_{\hat{t}} \frac{| \sum_K \lambda_K \langle \hat{t},1_K \rangle |}{|| \hat{t} ||_E}
\geq  \beta \LRp{\sum_K \mu(K) \lambda_K^2}^{1/2}\,.
\end{equation}
As in the previous analysis, we select for $\hat{t}$ the trace of Raviart-Thomas
interpolant $\bfpsi_h$ of $\bfpsi = R (\sum_K \lambda_K 1_K)$ where $R$ is the right-inverse
of the divergence operator constructed by Costabel and McIntosh. The only change compared
with the previous analysis, is the evaluation of norm of $\hat{t}_h$. For this, we need
to solve the local problems:
\begin{align}
((v,\bftau), (\delta v, \delta \bftau)_V &= \langle \hat{t}, \delta v \rangle_{\ptl K}
= \int_K \text{div} \bfpsi_h \, \delta v = \int_K \text{div} \bfpsi \, \delta v
\nonumber\\
&= \int_K \lambda_K \delta v
= \lambda_K (1_K,\delta v)_K  \quad \forall \delta v \in
H^1(K)\,\forall\delta\bftau\in\bfH(\text{div},K)\,.
\label{eq:local}
\end{align}
We need then an upper bound of the energy norm of $(v_h,\bftau_h)$:
$$
\LRp{\sum_K || (v,\bftau) ||_V^2}^{1/2}\,.
$$
Substituting $(v,\bftau)$ for $(\delta v,\delta\bftau)$ in~\eqref{eq:local}, we get,
\begin{equation}
|| (v,\bftau) ||_V^2 = \lambda_K (1_K, v_K)\,.
\end{equation}
If we have a robust stability estimate:
\begin{equation}
(1_K, v_K) \leq C \mu(K)^{1/2} || (v,\bftau) ||_K
\label{eq:robustEst1}
\end{equation}
(i.e. constant $C$ is independent of $\epsilon$), then
\begin{equation}
||  (v,\bftau) ||_V \leq C \mu(K)^{1/2} | \lambda_K |
\end{equation}
and, eventually as needed,
\begin{equation}
\sum_K || (v,\bftau) ||_V^2  \leq C^2 \sum_K \mu(K) \lambda_K^2\,,
\end{equation}
which leads to the robust estimate of inf-sup constant $\beta$. For example, it is sufficient if
\begin{equation}
|| v ||_{L^2(K)} \leq || (v,\bftau) ||_V \, .
\label{eq:robust_est1}
\end{equation}
Notice that the stability analysis with the energy norm was, in a sense, easier than
with the quotient norm. Only the divergence of the interpolant $\bfpsi_h$
enters~\eqref{eq:local}
and it coincides with the divergence of $\bfpsi$.

We arrive at a similar situation in the continuity estimate of
$$
\sum_K \lambda_K \langle \hat{t}, 1_K \rangle\,.
$$
Testing with $(1_K,{\bf 0})$ in the local problem~\eqref{eq:local_problem}, we obtain,
\begin{equation}
((v,\bftau),(1_K,{\bf 0}))_V = \langle \hat{t}, 1_K \rangle_{\ptl K}\,.
\label{eq:local2}
\end{equation}
If we have a robust estimate,
\begin{equation}
| ((v,\bftau),(1_K,{\bf 0}))_V | \leq C \mu(K)^{1/2} \, || (v,\bftau) ||_V\,,
\label{eq:robust_est2}
\end{equation}
then
\begin{equation}
| \sum_K \lambda_K \langle \hat{t}, 1_K \rangle | \leq C (\sum_K \mu(K) \lambda_K^2)^{1/2}
\, (\sum_K || (v,\bftau) ||_V^2)^{1/2}
= C (\sum_K \mu(K) \lambda_K^2)^{1/2}  || \hat{t} ||_E\,,
\end{equation}
as needed.

For instance, condition~\eqref{eq:robust_est2} will be satisfied if the test inner product
in~\eqref{eq:local2} reduces to the $L^2$ term only,
\begin{equation}
((v,\bftau),(1_K,{\bf 0}))_V = (v,1_K)_{L^2(K)} \: .
\end{equation}
With the robust stability and continuity constants for the mixed problem, the energy error
of solution $(u,\bfsig,\hat{u},\hat{t})$ (and Lagrange multipliers $\lambda_K$ as well)
is bounded robustly by the {\em best approximation error} of  $(u,\bfsig,\hat{u},\hat{t})$
measured in the energy norm. We arrive thus at the same situation as in the standard
DPG method.

@@

We now proceed to show that the robust norms we are using satisfy this
requirement.
Consider the inner product from \eqref{eq:localSolve}, with
$\alpha=1$. We wish to
verify condition \eqref{eq:robustEst1} with the norm derived from this inner
product on the right hand side. By Cauchy-Schwarz
\begin{equation}
   \int_K v\cdot
   1\leq\mu(K)^{1/2}\norm{v}_{L^2(K)}\leq\mu(K)^{1/2}\norm{(v,\btau)}_K\,,
   \label{eq:robustProof}
\end{equation}
where $\norm{(v,\btau)}_K$ is the norm derived from the inner product.
Condition~\eqref{eq:robust_est2} comes out the same since
$$|(
(v,\btau),(1_K,{\bf0}))| =
\sum_K|(1_K,v_K)|\leq\sum_K\mu(K)^{1/2}\norm{v,\btau}_K
$$
element-wise.

Now we wish the perform the same analysis for the modified inner product in
\eqref{eq:localSolveMod}. In this case,
condition~\eqref{eq:robustEst1} follows even more naturally as
\begin{equation}
   \int_K v\cdot1\leq\mu(K)^{1/2}\frac{1}{\mu(K)^{1/2}}\LRl{\int_Kv}
   \leq\norm{(v,\btau)}_K\,,
   \label{eq:robustProof2}
\end{equation}
where $\norm{(v,\btau)}$ now refers to the norm generated by inner product
\eqref{eq:localSolveMod}. Condition~\eqref{eq:robust_est2} follows by the same
reasoning.
